install.packages('rmarkdown')

%then open a markdown file
title: "test"
author: "Carlotta Zanetti"
date: "2024-01-08"
output: html_document
---

%press knit to see the result of the code
%then delete everything and write:

# RMarkDown file! %it's the title
Here you can write something

%now let's add the code
```{r, eval=T}
2+2=4
```
%eval=T means that u print out the output

%let's try to use proper code:
```{r, eval=T}
library(imageRy)
```
```{r, eval=T}
im.list()
```
```{r, eval=T}
mato1992 <-im.import("matogrosso_l5_1992219_lrg.jpg" )
```
```{r, eval=T}
mato2006 <-im.import("matogrosso_ast_2006209_lrg.jpg")
```
in order to have the infrmation about the object, just type the name of the object
mato1992
##RGB

```{r, eval=T}
im.plotRGB(mato1992, r=2, g=1, b=3)
```
This is with true colors. If you want to use the NIR on top of the green channel, you will have to write g=8, instead of g=2
```{r, eval=T}
im.plotRGB(mato1992, r=2, g=1, b=3)
```
If the programme gives you warning, it's because of eco.
So you can turn it down
```{r, eval=T, warning=F}
mato1992 <-im.import("matogrosso_l5_1992219_lrg.jpg" )
```
```{r, eval=T, warning=F}
mato2006 <-im.import("matogrosso_ast_2006209_lrg.jpg")
```
Now let's plot 4 images
```{r, eval=T}
par(mfrow=c(2,2))
im.plotRGB(mato1992, 1,2,3)
im.plotRGB(mato1992, 2,1,3)
im.plotRGB(mato1992, 3,2,1)
im.plotRGB(mato1992, 1,3,1)
```

```{r, eval=T}
im.list()
```

```{r, eval=T}
mato2006 <-im.import("matogrosso_ast_2006209_lrg.jpg")
```

```{r, eval=T}
im.plotRGB(mato2006, 2, 1, 3)
```

Calculating spectral indices:
```{r, eval=T}
library(terra)
library(viridis)
dvi <- mato1992[[1]]-mato1992[[2]]
viridisc <- colorRampPalette(viridis(7))(255)
plot(dvi,col=viridisc)
```
Calculating variability
```{r, eval=T}
sd5 <- focal(mato1992[[1]], matrix(1/25))
plot(dvi,col=viridisc)
```
